A boundary-integral framework to simulate \\ viscous erosion of a porous medium

### 2018-04-12 docs/methods/erosion_methods.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large-scale land formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, down to human-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and the related processes of dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow'~\cite{ristroph2018sculpting} have been examined in the context of a single body~\cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shape evolution is coupled through the flow~\cite{hewett2017evolution}. From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.
% in particular close-contacts?

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.80 \textwidth]{./figs/50bod.pdf}
\caption{\label{fig:50bodies} A time lapse of 50 bodies eroding in a viscous flow.  The boundary conditions are a constant pressure drop that creates a flow from the left to right.  The rate of erosion is proportional to the shear stress which is equivalent to the vorticity evaluated on the bodies.  The colors represent the vorticity, so regions with large vorticity (in magnitude) are regions where the flow will erode the fastest.}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction\label{s:intro}}
Flow-induced erosion deteriorates and reshapes solid material over a range of scales found in nature, from large land formations~\cite{han1969, Rothman2012, coh-dev-sey-yi-szy-rot2015}, to centimeter-scale features and patterns~\cite{daerr2003erosion}. Though less visible, these same forces are working at the very smallest scales, slowly deteriorating the individual constituents of porous media (e.g.~solid, sand, or clay) or biological structures like plaque and biofilms~\cite{bear2013dynamics, pic-van-hei2000, sha2002, gro-gij-van-fer-hat-van-yua-wen2007}. Motivated by such examples, this paper develops numerical methods to simulate fluid-mechanical erosion of multiple solid bodies in a Stokes flow---the most relevant regime for groundwater and the aforementioned bio-fluid applications.

Recent work in the high-Reynolds-number regime has highlighted the importance of a shape-flow feedback that occurs during erosion~\cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, hewett2017evolution, moore2017riemann} and during related processes such as dissolution and melting~\cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, hewett2017pear, claudin2017dissolution}.  In these cases, the flow alters the morphology of the immersed structures, which in turn modifies the surrounding flow, and so on. Such processes of `sculpting with flow'~\cite{ristroph2018sculpting} have been examined in the context of a single body~\cite{ris-moo-chi-she-zha2012, moore2017riemann} and, more recently, multiple bodies whose shape evolution is coupled through the flow~\cite{hewett2017evolution}. From this viewpoint, the example of an eroding porous medium presents new opportunities and challenges. First, since the medium is composed of many individual constituents, multi-body interactions must be considered. Second, the long-range effects inherent to Stokes flow implies that the shape-evolution of bodies are even more strongly coupled than in any other regime.
% in particular close-contacts?

@@

%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[htbp]
\begin{center}
\includegraphics[width = 0.80 \textwidth]{./figs/50bod.pdf}
\caption{\label{fig:50bodies} Simulation of 50 bodies eroding in Stokes flow under the action of fluid-mechanical shear. The flow is horizontal (left to right) and the 6 snapshots are evenly spaced in time. Color represents vorticity, which provides a convenient way to visualize local shear rates. Erosion not only diminishes the size of the bodies but also alters their shapes considerably. A few well-defined channels develop in the space between bodies. Several of the bodies vanish in finite time and the simulation continues.
}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Figure generated from data file 50circ512p5, 
%which uses 50circ512.in with epsfac = 30, sigfac = 10, dt = 1e-4, fixpdrop = 1
